---
title: "CV Paul Beard"
format: pdf
editor: visual
execute: 
  echo: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)

data.loc <- "./Data"
data.filename <- "Work History Summary.xlsx"

# Read in data
data.filepath <- paste(data.loc,"/",data.filename,sep="")
read.all.excel.sheets <- function(filename, tibble = FALSE) {
  sheets <- readxl::excel_sheets(filename)
  x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
  if(!tibble) x <- lapply(x, as.data.frame)
  names(x) <- sheets
  x
}
cvdata <- read.all.excel.sheets(data.filepath)

```

## **Paul Beard FIA**

**Summary**

-   I am a high performing Fellow of the Institute of Actuaries working in the life insurance industry for 10 years, with 5 years post qualification. I have a diverse range of work experience with my main areas of speciality being:

    -   With Profits Management

    -   Internal Model Risk Calibration

    -   Commercial Real Estate Loans 

<!-- -->

-   I am proficient across a variety of platforms, with emphasis on Excel and R (see Figure 1). I can develop complex models; build self-checks; debug established code; and peer review work. I am also capable of methodology development, and have done this across a variety of areas through my career. I am meticulous and am able to think through any complexities encountered. I am technically minded and am capable of finding solutions to novel problems.

<!-- -->

-   I have a fantastic track record of performance, both academically (see Table 1) and in a work environment (see Figure 2).

**Qualifications**

**Table 1: Qualification History**

| Qualification | Awarded By              | Grade            | Subject(s)                   | Date     |
|---------------|---------------|---------------|---------------|---------------|
| Fellowship    | IFoA                    |                  | SA2, ST2, ST6                | Dec 2017 |
| Bachelors     | University of Cambridge | 1                | Natural Sciences (Chemistry) | Jul 2011 |
| Masters       | University of Cambridge | 2.i              | Natural Sciences (Chemistry) | Jul 2011 |
| A Level       |                         | A (x5)           | Maths, Further Maths, Other  | Aug 2007 |
| GCSE          |                         | A\* (x8); A (x3) | Maths, English, Other        | Aug 2005 |

Other achievements include:

-   One of top 5 scores in the UK for GCSE Maths (awarded by AQA in Aug 2005)

-   IFoA Certificate in Data Science (awarded by IFoA in Dec 2020)

**Work Experience**

**Phoenix Group (Apr 2023 -  Feb 2024)** - Contractor 

BPA Assumptions Architecture (April 2023 - Feb 2024)

-   Phoenix uses MG-Alfa as their main valuation platform. The tables used by MG-Alfa are created through a set of complex spreadsheets. In this role I developed the spreadsheets that fed BPA assumptions into the MG-Alfa capital model. This was part of a project to adapt the model to an increasing BPA portfolio. 

-   These spreadsheets contained a great deal of complexity after years of projects building piecemeal over the top. Therefore, as well as developing I often performed deep dives to restructure the spreadsheets to make the calculation clearer, more efficient and transparent.

-   Created a sophisticated comparison tool in R which could compare two complex assumptions spreadsheets. By replicating the tenets of how MG-Alfa reads table structures, the tool can produce an intelligent breakdown of the changes that were made to the assumptions tables. The tool was packaged up and rolled out to assist other actuaries working in the assumptions space.

**Royal London (Sep 2012 -  Apr 2023)** - Permanent

Credit Quality Assessor of Commercial Real Estate Loans (CRELs) (August  2021 - April 2023)

-   Royal London looked to expand the illiquid asset classes it invests in as part of its Matching Adjustment portfolio. This entailed the inclusion of CRELs. In this role I developed an internal credit quality assessment process (ICQA) for CRELs. As this asset class was new to the company, it required a lot of development from ground level.

-   As part of the role I created a complex stochastic model in R which captures the intricate cash flow structure of CRELs. The model simulates defaults across multiple correlated tenants over many time periods, and evaluates the outcomes of the loans to assess the credit quality. 

-   After developing the ICQA process for CRELs, I developed the internal model risk calibration for CREL risk. This combined my experience of CRELs with my experience of internal model calibration. As per the ICQA methodology the I developed most of the framework. 

Actuarial Student Overseer (June 2018 - November 2021)

-   Volunteered as coordinator and advisor to over 60 students sitting exams at Royal London. This role has required strong communication and leadership skills. I made sure to always be approachable to the students, and keep open paths of dialogue. 

Internal Model Calibration (March 2018 - August 2021)

-   I worked on the development of the company’s Internal Model Application Process (IMAP). This covered a range of insurance risk type calibrations where I compiled reports on Longevity Trend, Longevity Level, GAO take-up, Persistency Level, Mortality and Morbidity risk. Some of these reports were very significant in size (\>100,000 words). I have also completed market risk calibrations including Commodity, Property and Market-to-market Dependencies.

-   The role required me to adapt many different processes, some that were developed by staff no longer present at the company. This required being able to understand methods with incomplete information and work autonomously, with little support. Have often had to make expert judgements. 

-   Answered questions from both internal validators and the PRA about our IMAP methodology in a short period of time. Resolved validation findings in a timely manner.

-   I have led many workshops to co-ordinate the opinions of experts in their respective fields, and form consensus on the assumptions and methodology used in the internal model calibrations.

With Profits Management (Sept 2012 – March 2018)

-   Within this role I carried out regular analysis on bonus supportability for With Profits business and became a subject matter expert and led a team of actuarial staff to complete year end final bonus reports.

-   Involved in many with profits management tasks including (but not limited to) bonus setting, projections, run off planning, estate distribution and remediations.

-   I have been involved in many bespoke projects, the most material being:

    -   Guaranteed Annuity Option (GAO) compromise scheme, where I priced values of GAOs and proposed appropriate buyout offers under the scheme. This was done to the satisfaction of both an Independent Actuary and With Profits Actuary. 

    -   Surrender value changes, where I developed novel spreadsheet models capable of calculating factors that would align surrender values.

    -   Prophet to MG-Alfa model reconciliation, where I reconciled asset share outputs from a newly developed MG-Alfa model, with the legacy Prophet model. This involved debugging the cash flow model calculations.

**Software Experience**

**Figure 1: Summary of Modelling Experience**

```{r}
#| label: software-experience

experience <- cvdata$Software

e.long <- experience %>%
  gather(key = year, value = hours, -c("Platform", "Totes")) %>%
  filter(!is.na(hours))

# get order of platforms in terms of total experience. These are p.levels and used to order platforms within plot at the end
p.levels <- e.long %>%
  group_by(Platform) %>%
  summarise(totes = sum(hours, na.rm = T)) %>%
  arrange(desc(totes)) %>%
  .$Platform %>%
  rev()

# note that when chunk.size is larger, the amount of hours in each row drops
# if amount of hours < 1 for a row, log10 will be negative. This can lead to graph going funky as some of the initial numbers are negative, and the total  of positive values is larger because sum is still the same. Graph only shows positive values so get weird stuff.
years <- e.long$year %>%  unique() %>% as.numeric() %>% sort()

e.list <- e.long %>% arrange(Platform, year) %>% split(~Platform)

e.cumloghrs <- purrr::map(e.list,
                          .f = ~ c(
                            tail(.$hours, 1) %>% log10() %>% replace(is.infinite(.), 0),
                            .$hours %>%
                              rev() %>%
                              cumsum() %>%
                              log10() %>%
                              replace(is.infinite(.), 0) %>%
                              diff(1)
                          )
)

ggdata <- e.cumloghrs %>%
  bind_cols() %>%
  tibble(year = rev(years), .) %>%
  pivot_longer(names_to = "Platform", values_to = "hours", -year) %>%
  arrange(desc(year))

#check no negative values
if(min(ggdata$hours)<0) {stop("graph will distort because log hours include negative values")}

# set up x axis parameters
x.max <- ggdata %>%
  group_by(Platform) %>%
  summarise(total = sum(hours)) %>%
  .$total %>%
  max()
x.min <- 1
x.breaks <- c(10,100,1000,10000) %>% log10()
x.labels <- 10^x.breaks

# set up mid-year point for colour scale
mid.year <-  min(years) + 0.5*(max(years) - min(years)) 

# set platform as factors and with levels in order of size
ggdata$Platform <- factor(ggdata$Platform, levels = p.levels)

# plot of software experience (final output)
ggplot(ggdata, aes(
  x = hours, y = Platform,
  fill = year
)) +
  geom_col(width=0.62) +
  scale_fill_gradient2(
    name = "Year",
    breaks = scales::pretty_breaks(),
    low = "grey97",
    mid = "steelblue1",
    midpoint = mid.year,
    high = "steelblue4"
  ) +
  coord_cartesian(xlim = c(x.min, x.max)) +
  scale_x_continuous(
    name = "Estimated Cumulative Experience (Hours)",
    breaks = x.breaks,
    minor_breaks = NULL,
    labels = scales::comma(x.labels)
  ) +
  labs(caption = "**please note the scale is logarithmic") +
  theme_classic() + 
  theme(
    panel.grid.major.x = element_line()
  )
```

**Performance**

Whilst I was employed by Royal London during years 2013 - 2020 I was assigned a performance rating each year, as were all staff. Employees were placed in one of four bands which represent different percentile ranges of performance relative to colleagues. Those bands, and the amount of times I was placed in each band over the 8 year period, are: Improvement Required (0); Good (3); Strong (4) and Exceptional (1). 

Below I have mapped my historical ratings to a probability density of performance across all percentiles. The density is calculated by the relative frequency of each rating awarded to me, divided over the percentiles of employee performance that rating represents. I have also included the probability density you would expect from an actuary selected randomly from Royal London, for comparison purposes.

**Figure 2: Performance rating density whilst at Royal London**

```{r}
#| label: RL-Performance
#| 
actuary.levels <- c("Random Actuary", "Paul Beard")
grade.levels <- cvdata$`RL Performance`$grade

data <- cvdata$`RL Performance` %>%
  mutate("Paul Beard" = count / (maxpercentile - minpercentile) / sum(count)) %>%
  mutate("Random Actuary" = 1) %>%
  pivot_longer(cols = c("Paul Beard", "Random Actuary"), names_to = "actuary", values_to = "density") %>%
  select(grade, minpercentile, maxpercentile, density, actuary)

data$grade <- factor(data$grade, levels = grade.levels)
data$actuary <- factor(data$actuary, levels = actuary.levels)

ggplot(data, aes(ymin = 0)) +
  geom_rect(
    aes(
      xmin = minpercentile,
      xmax = maxpercentile,
      ymax = density,
      fill = grade,
    ),
    colour = "white"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = c(seq(0, 1, 0.2)), labels = scales::percent(seq(0, 1, 0.2))) +
  scale_y_continuous(breaks = c(seq(0, 3, 0.5))) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "light grey")
  ) +
  scale_fill_manual(name = "Performance Rating", values = c(
    rgb(0.2, 0.2, .8, 0.25),
    rgb(0.2, 0.2, .8, 0.45),
    rgb(0.2, 0.2, .8, 0.65),
    rgb(0.2, 0.2, .8, 0.9)
  )) +
  labs(
    x = "Percentile of Performance",
    y = "Probability density"
  ) +
  facet_wrap(~actuary)
```

**Other Projects**

I write and publish articles applying actuarial tools to publicly available data. These include:

-   A mortality investigation of Coronation Street

-   An internal model risk calibration of cryptocurrency risk

-   An analysis of trends in IFoA exam pass rates

The articles can be found below, along with the R code used to produce the analyses.

Articles: <https://www.linkedin.com/in/paul-beard-78a420172/recent-activity/posts/> 

Code: <https://github.com/paulbeardactuarial> 
